\specialhead{ACKNOWLEDGMENT}

I would first of all like to thank my parents, Arindam Chowdhury and Babli Chowdhury for everything. I am here because of them and anything I do will never be enough to repay the sacrifices that they made for me. It is an honour to be their son. They have supported me in my journey through thick and thin and I am very grateful for everything that they have done for me.

I would like to specially  thank my adviser, Prof. B{\"u}lent Yener for guiding me throughout my Ph.D. His advice about everything, starting from the projects that we have worked on together, to how to navigate life in the academic world, has proved to be invaluable. It has been
a wonderful learning experience for me to have had the opportunity to work with
him these past 5 years. He has been a role model for me and many others with his professionalism, openness, caring attitude, and big heart. 
I would  like to thank my committee members - Prof. Malik Magdon-Ismail, Prof. Charles Stewart and Dr. Alberto Santamaria-Pang for helping me with the thesis and guiding me in the right direction.  I am grateful that I have had the opportunity to work with  Prof.  Magdon-Ismail, Prof. Dan Lewis and Dr. Santamaria-Pang, and learn immensely from them.


I would also like to show my gratitude to my friends and colleagues here at RPI. It was a lot of fun discussing about research and life with my friends and colleagues (Kshiteesh Hegde, Maksim Tsikhanovich, Nivas Nambirajan, Nimit Dhulekar, Haidar Khan, Elizabeth Kautz, Daniel Park, Gilna Samuel, John Postl, Anik Saha).
I would also like to specially thank my roommates Atriya Sen and Partha Sarathi Mukherjee who have proved to be my family away from home.


I would also like to acknowledge the staff at Lally - Christine Coonrad, the late Terry Hayden (who we all miss very much), Mike (the environmental specialist), Tracy Hoffman and others. I would like to thank you for making our work life easy. 

This section would be incomplete without mentioning certain establishments like First Choice Carribean, Red and Blue and Snowman's ice cream. They have been a source of joy and nourishment throughout the years of my Ph.D.